---
title: "Reproducibility and transparency in interpretive corpus pragmatics - Part 2: data analysis"
author: "Martin Schweinberger and Michael Haugh"
date: "June 12, 2023"
output: html_document
---

# Preparation

In a first step, we install the packages that we need for the analysis.

```{r setup1, eval = F, message=F, warning=F}
install.packages("rlang")
install.packages("tidyverse")
install.packages("quanteda")
install.packages("here")
install.packages("openxlsx")
install.packages("knitr")
```


After the installation, we load or activate the packages.

```{r setup2, message=F, warning=F}
library(tidyverse)
library(quanteda)
library(here)
library(openxlsx)
library(knitr)
```


# Data Exploration and Analysis

We now load the manually annotated data.

```{r ufor_ann, warning=F, message=F}
ufor_ann <- openxlsx::read.xlsx(here::here("tables", "ufors_annotated.xlsx"), sheet = 1)
# inspect
ufor_ann %>%
  dplyr::filter(corpus == "The La Trobe Corpus of Spoken Australian English") %>%
  head() %>%
  knitr::kable()
```

You can use and edit the code chunk below to inspect other instances of utterance-final *or* by changing the identifier of the instance, e.g., from `instance 1` (which is the default below) to `instance 51` (overall, there are 98).

```{r}
# inspect
ufor_ann %>%
  dplyr::filter(hit == "instance 1") %>%
  knitr::kable()
```


We will now generate an overview table showing us how frequent different combinations are in the raw data.

```{rufor_anntb, warning=F, message=F}
ufor_ann %>%
  dplyr::group_by(hit) %>%
  tidyr::fill(action.type, .direction = "updown") %>%
  tidyr::fill(question.type, .direction = "updown") %>%
  tidyr::fill(response.polarity, .direction = "updown") %>%
  tidyr::fill(`explicit-inferred`, .direction = "updown") %>%
  tidyr::fill(response.type, .direction = "updown") %>%
  dplyr::filter(context == "hit") %>%
  group_by(action.type, question.type, response.polarity, `explicit-inferred`, response.type) %>% 
  dplyr::summarise(Frequency = n()) %>%
  dplyr::arrange(-Frequency) %>%
  knitr::kable()
```



> It is important to note that there are false positives in the data, i.e. instances that do not really represent instances of utterance-final *or*. Hence, we will remove all instances representing false positives but also non-canonical uses of utterance-final *or* from the data as the analysis will focus on canonical uses of utterance-final *or*. Cononical instances are where the *or* is part of a ploar question.

## Canonical (Q-P) Instances

```{r cantb, warning=F, message=F}
# inspect
ufor_can <- ufor_ann %>%
  dplyr::group_by(hit) %>%
  tidyr::fill(action.type, .direction = "updown") %>%
  tidyr::fill(question.type, .direction = "updown") %>%
  tidyr::fill(response.polarity, .direction = "updown") %>%
  tidyr::fill(`explicit-inferred`, .direction = "updown") %>%
  tidyr::fill(response.type, .direction = "updown") %>%
  # filter canonical instances
  dplyr::filter(action.type == "Q" & question.type == "P")
# inspect
knitr::kable(head(ufor_can))
```


We check how many instances of utterance-final *or* are left.

```{r}
length(names(table(ufor_can$hit)))
```

We are left with 57 canonical instances of utterance-final *or* (i.e. where the utterance containing utterance-final *or* is an information seeking question (Q) and a polar question (P).


We will now check, what instances are left in the data.

```{r}
names(table(ufor_can$hit))
```


You can use and edit the code chunk below to inspect other instances of utterance-final *or* by changing the identifier of the instance, e.g., from `instance 1` (which is the default below) to `instance 9` (overall, there are 57).

```{r}
# inspect
ufor_can %>%
  dplyr::filter(hit == "instance 1") %>%
  knitr::kable()
```



We will now generate an overview table showing us how frequent different combinations are in the canonical data.

```{r ufor_camoverview, warning=F, message=F}
ufor_can %>%
  dplyr::filter(context == "hit") %>%
  group_by(action.type, question.type, response.polarity, `explicit-inferred`, response.type) %>% 
  dplyr::summarise(Frequency = n()) %>%
  dplyr::arrange(-Frequency) %>%
  knitr::kable()
```

## Canonical with Y or N response

We now want to check the instances where the canonical sequence has received either a positive *yes* [Y] or a negative *no* [N] responses.

```{r}
ufor_can_yn <- ufor_can %>%
  dplyr::filter(response.polarity == "Y" | response.polarity == "N")
# inspect
knitr::kable(head(ufor_can_yn))
```

Again, we will now check, how many instances are left in the data.

```{r}
length(names(table(ufor_can_yn$hit)))
```
We see that there are 46 instances left of canonical sequences where the response is positive or negative.

We will now check, what instances are left in the data.

```{r}
names(table(ufor_can_yn$hit))
```


You can use and edit the code chunk below to inspect other instances of utterance-final *or* with positive and negative responses by changing the identifier of the instance, e.g., from `instance 1` (which is the default below) to `instance 82` (overall, there are 46).

```{r ufor_can_yn_overview, message=F, warning=F}
# inspect
ufor_can_yn %>%
  dplyr::filter(hit == "instance 1") %>%
  knitr::kable()
```


We will now generate an overview table showing us how frequent different combinations are in the canonical data  with positive and negative responses.

```{r ufor_can_yn_overview2, message=F, warning=F}
ufor_can_yn %>%
  dplyr::filter(context == "hit") %>%
  group_by(action.type, question.type, response.polarity, `explicit-inferred`, response.type) %>% 
  dplyr::summarise(Frequency = n()) %>%
  dplyr::arrange(-Frequency) %>%
  knitr::kable()
```

## Canonical with explicit Y or N response

We now want to check the instances where the canonical sequence has received either an **explicit** positive *yes* [Y] or negative *no* [N] responses.

```{r}
ufor_can_eyn <- ufor_can_yn %>%
  dplyr::filter(`explicit-inferred` == "E")
# inspect
knitr::kable(head(ufor_can_eyn))
```

Again, we will now check, how many instances are left in the data.

```{r}
length(names(table(ufor_can_eyn$hit)))
```

Also, we check, what instances are left in the data.

```{r}
names(table(ufor_can_eyn$hit))
```

You can use and edit the code chunk below to inspect other instances of utterance-final *or* with positive and negative responses by changing the identifier of the instance, e.g., from `instance 1` (which is the default below) to `instance 75` (overall, there are 37).

```{r}
# inspect
ufor_can_eyn %>%
  dplyr::filter(hit == "instance 1") %>%
  knitr::kable()
```

We will now generate an overview table showing us how frequent different combinations are in the canonical data  with explicit positive and negative responses.

```{r ufor_can_yn_overview5, message=F, warning=F}
ufor_can_eyn %>%
  dplyr::filter(context == "hit") %>%
  group_by(action.type, question.type, response.polarity, `explicit-inferred`, response.type) %>% 
  dplyr::summarise(Frequency = n()) %>%
  dplyr::arrange(-Frequency) %>%
  knitr::kable()
```


# Outro

```{r}
sessionInfo()
```


