---
title: "Reproducibility and transparency in interpretive corpus pragmatics - Part 1: data extration and processing"
author: "Martin Schweinberger and Michael Haugh"
date: "June 12, 2023"
output: html_document
---

# Preparation

In a first step, we install the packages that we need for the analysis.

```{r setup1, eval = F, message=F, warning=F}
install.packages("rlang")
install.packages("tidyverse")
install.packages("quanteda")
install.packages("here")
install.packages("openxlsx")
install.packages("knitr")
```


After the installation, we load or activate the packages.

```{r setup2, message=F, warning=F}
library(tidyverse)
library(quanteda)
library(here)
library(openxlsx)
library(knitr)
```

# Loading and Processing

For this analysis, we will use dta from the following corpora:


* [Australian Radio Talkback](https://www.ausnc.org.au/corpora/art)

* [Griffith Corpus of Spoken Australian English](https://www.ausnc.org.au/corpora/gcsause)

* [Monash corpus](https://www.ausnc.org.au/corpora/monash)

* [The La Trobe Corpus of Spoken Australian English](https://www.ausnc.org.au/corpora/latrobecsause)

## Loading data

The corpora were downloaded and stored in a directory (folder) called `data`. To load the data, we define the paths to the files containing the transcripts (which are located in the `data` folder in the specific sub-directories for the corpora).

```{r}
fart <- list.files(here::here("data", "Australian Radio Talkback/files/Raw"), 
                        pattern = ".txt", full.names = T)
fgri <- list.files(here::here("data", "Griffith Corpus of Spoken Australian English/files/Raw"), 
                        pattern = ".txt", full.names = T)
fmon <- list.files(here::here("data", "Monash/files/Text"), 
                        pattern = ".txt", full.names = T)
flat <- list.files(here::here("data", "The La Trobe Corpus of Spoken Australian English/files/Raw"), 
                        pattern = ".txt", full.names = T)
```


We now check if we have the paths to the data by inspecting the first six paths of files in the *Australian Radio Talkback* corpus.



```{r}
# inspect
knitr::kable(head(fart))
```

We now proceed by loading and processing (cleaning) the data.

### Load ART

We start with the  content of the *Australian Radio Talkback* corpus (art).

```{r lart, message=F, warning=F}
# load raw content
vart <- sapply(fart, function(x){
  x <- readLines(x)
  x <- x[x != ""]
  })
# unlist the object containign the corpus data
arttext <- unlist(vart)
# collapse into a data frame
artdf <- data.frame(names(arttext), names(arttext),arttext) %>%
  # rename columns
  dplyr::rename(corpus = colnames(.)[1],
                file = colnames(.)[2],
                text = colnames(.)[3]) %>%
  # create new columns containing corpus, file, and speaker information as well as a column with clean content
  dplyr::mutate(
    # extract corpus name
    corpus = stringr::str_replace_all(corpus, ".*data/(.*?)/.*", "\\1"),
    # extract file name
    file = stringr::str_replace_all(file, ".*Raw/(.*?)-raw.*", "\\1"),
    # extract speaker
    speaker = stringr::str_replace_all(text, "\\[(.*?)\\].*", "\\1"),
    # extract clean transcript
    textclean = stringr::str_remove_all(text, ".*?\\]"),
    # remove superfluous white spaces
    textclean = stringr::str_squish(textclean))
# remove row names
rownames(artdf) <- NULL
# inspect
knitr::kable(head(artdf))
```


### Load GRI

We continue with the content of the files of the *Griffith Corpus of Spoken Australian English* (gri).

```{r lgri,  message=F, warning=F}
vgri <- sapply(fgri, function(x){
  x <- readLines(x, encoding = "UTF-8")
  x <- x[x != ""]
  x <- x[!stringr::str_detect(x, "\\|.*\\|")]
  x <- paste0(x, collapse = " ")
  x <- stringr::str_split(stringr::str_replace_all(x, "( [A-Z]:)", "qwertz\\1"), "qwertz")
  x <- unlist(x)
  x <- stringr::str_squish(x)
})
gritext <- unlist(vgri)
# collapse into df
gridf <- data.frame(names(gritext), names(gritext),gritext) %>%
  dplyr::rename(corpus = colnames(.)[1],
                file = colnames(.)[2],
                text = colnames(.)[3]) %>%
  dplyr::mutate(corpus = stringr::str_replace_all(corpus, ".*data/(.*?)/.*", "\\1"),
                file = stringr::str_replace_all(file, ".*Raw/(.*?)-raw.*", "\\1"),
                speaker = stringr::str_remove_all(text, ":.*"),
                speaker = stringr::str_remove_all(speaker, "\\W.*\\W"),
                speaker = stringr::str_remove_all(speaker, "[^[:alpha:]]"),
                speaker = stringr::str_remove_all(speaker, "[a-z]"),
                textclean = stringr::str_remove_all(text, "<.*?>"),
                textclean = stringr::str_remove_all(textclean, "(.*?)"),
                textclean = stringr::str_remove(textclean, "[0-9]{0,} {0,}[A-Z]{1,}:"),
                textclean = stringr::str_remove_all(textclean, "[^[:alpha:]’ ]"),
                textclean = stringr::str_squish(textclean))
rownames(gridf) <- NULL
# inspect
knitr::kable(head(gridf))
```


### Load MON

We continue with the content of the files of the *Monash corpus* (mon)

```{r lmon,  message=F, warning=F}
vmon <- sapply(fmon, function(x){
  x <- readLines(x, encoding = "UTF-8")
  x <- x[x != ""]
  x <- paste0(x, collapse = "qwertz") %>%
  stringr::str_remove_all("qwertz    ") %>%
  stringr::str_split("qwertz") %>%
  unlist() %>%
  stringr::str_squish()
  })
# unlist
montext <- unlist(vmon)
# collapse into df
mondf <- data.frame(names(montext), names(montext), montext) %>%
  dplyr::rename(corpus = colnames(.)[1],
                file = colnames(.)[2],
                text = colnames(.)[3]) %>%
  dplyr::mutate(corpus = stringr::str_replace_all(corpus, ".*data/(.*?)/.*", "\\1"),
                file = stringr::str_replace_all(file, ".*Text/(.*?).txt", "\\1"),
                speaker = paste0("NA"),
                textclean = stringr::str_squish(text))
rownames(mondf) <- NULL
# inspect
knitr::kable(head(mondf))
```


### Load LAT

We continue with the content of the files of the *The La Trobe Corpus of Spoken Australian English* (lat)

```{r llat,  message=F, warning=F}
vlat <- sapply(flat, function(x){
  x <- readLines(x, encoding = "UTF-8")
  x <- x[x != ""]
  })
lattext <- unlist(vlat)
# collapse into df
latdf <- data.frame(names(lattext), names(lattext),lattext) %>%
  dplyr::rename(corpus = colnames(.)[1],
                file = colnames(.)[2],
                text = colnames(.)[3]) %>%
  dplyr::mutate(corpus = stringr::str_replace_all(corpus, ".*data/(.*?)/.*", "\\1"),
                file = stringr::str_replace_all(file, ".*Raw/(.*?)-raw.*", "\\1"),
                speaker = stringr::str_remove_all(text, ":.*"),
                speaker = stringr::str_remove_all(speaker, "\\W.*\\W"),
                speaker = stringr::str_remove_all(speaker, "[^[:alpha:]]"),
                speaker = stringr::str_remove_all(speaker, "[a-z]"),
                textclean = stringr::str_remove_all(text, "^[A-Z]{1,}:{0,1}"),
                textclean = stringr::str_squish(textclean))
rownames(latdf) <- NULL
# inspect
knitr::kable(head(latdf))
```


# Collapse data sources

We now combine the corpora into a single data frame called *oz*.

```{r}
oz <- rbind(artdf, gridf, mondf, latdf)
# inspect
knitr::kable(head(oz))
```

# Extract utterance-final or

In a next step, we extract utterances with utterance final *or*. We determine this by checking if a string (utterance) ends with the sequence *or* but we allow for another words to come after the or if it has up to three chacraters (e.g., "... or uhm?").

```{r}
ufor <- oz %>%
  dplyr::mutate(ufor = ifelse(stringr::str_detect(textclean, " or {0,}.{0,3}$"), 1, 0)) %>%
  dplyr::filter(ufor == 1)
# inspect
knitr::kable(head(ufor$textclean))
```

In a next step,w e want to extract concordances (keywords-in-context) of potential hits (utterance-final *or*). The context should be two utterances preceding the utterance with utterance-final *or* and two utterances following the instance of utterance-final *or*.


```{r ufors, message=F, warning=F}
inds = which(stringr::str_detect(oz$textclean, " or {0,}.{0,3}$"))
# We use lapply() to get all rows for all indices, result is a list
rows <- lapply(inds, function(x) (x-2):(x+2))
# With unlist() you get all relevant rows
ufors <- oz[unlist(rows),]
# insepct
knitr::kable(head(ufors))
```

We now generate a table with the instances of utterance-final *or* and the preceding as well as subsequent utterances and save the data to out computer for the manual annotation of the functions of utterance-final *or*.


```{r warning=F, message=F}
# label instances
nhits <- sapply(rows, function(x){ length(x) })
nints <- 1:length(rows)
labs <- rep(paste0("instance ", nints), each = nhits)
# label context
contlabs <- rep(c("pre2", "pre1", "hit", "post1", "post2"), length(rows))
# add to kwics
ufors <- ufors %>%
  dplyr::mutate(hit = labs,
                context = contlabs) %>%
  dplyr::select(-speaker, -textclean) %>%
  dplyr::relocate(corpus, file, hit, context, text)
```

The data frame now contains five lines for each instance: *pre2*, *pre1*, *hit*, *post1*, and *post2*. The instance of utterance-final *or* is shown in the row labeled as *hit*.

```{r uforstb, warning=F, message=F}
# inspect
knitr::kable(head(ufors))
```


# Save data to disc

We now save the data so that we can annotate and code the data manually in a spreadsheet software (MS Excel).

```{r eval = F}
# save
write.xlsx(ufors, here::here("tables", "ufors.xlsx"), sheetName = "Sheet1", 
           col.names = TRUE, 
           row.names = TRUE, 
           append = FALSE)
```

# Manula Annotation (Coding)

The data (ufors) were annotated manually in a spreadsheet software (MS Excel). 

## UF-or data annotation scheme

The annotation scheme used to code individual instances of utterance-final *or* is provided below. Each instance was inspected and annotated with regard to the categories shown below.

| **Question type**				 | Action type**			       | **Response polarity**		    | **Responses to alternative questions**  | **Explicit-implicit (dis)confirmation**	  | **Response type**			   | **Annotator comment**					       | **Turn-initial particle**
|------------------------------------------------|---------------------------------------------|------------------------------------|-----------------------------------------|-------------------------------------------|----------------------------------------|-------------------------------------------------------------------|-------------------------------------
| Polar question [P]				 | Information-seeking question [Q]	       | Responses to polar questions:	    | First alternative [A]		      | Explicit (yes/no, direct repeat) [E]	  | Type-conforming (yes/no; A or B) [TC]  | Points to potentially follow-up; notes on more complex cases      | um, uh, ah, oh, well, look, mm, no, 
| Alternative question [A]			 | Request (permission-seeking question) [R]   | Yes [Y]			    | Second alternative [B]		      | Implicit [I]				  | Non-type-conforming
| Q-word question [Q]				 | Assertion [A]			       | No [N]				    | Neither [N]
| False positive (i.e. not a question) [FP]	 | Suggestion [S]			       | Yes-no [YN]
|------------------------------------------------|---------------------------------------------|------------------------------------|-----------------------------------------|-------------------------------------------|----------------------------------------|-------------------------------------------------------------------|-------------------------------------											       | Non-answer [NA]


We now annotate the data manually in a spreadsheet software (MS Excel) according to the annotation/coding schema provided above.


# Outro

```{r}
sessionInfo()
```


